Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-27 paper.tex
Fix overfull hboxes

In many proof assistants, simple \enquote*{white-box} automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's~\cite{Isabelle} and Rocq's~\cite{Coq} auto~\cite{IsabelleAuto}, ACL2's \enquote{waterfall}~\cite{ACL2}, various PVS~\cite{PVS-tutorial,PVS} tactics, Lean's~\cite{Lean4} Aesop~\cite{Aesop}, etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.
A typical rule is the $∧$-introduction lemma $A → B → A ∧ B$, which, when applied backwards, reduces a goal $Γ ⊢ A ∧ B$ to subgoals $Γ ⊢ A$ and $Γ ⊢ B$.

@@

\para{White-Box Tactics.}
Among white-box proof automation tactics, to our knowledge only ACL2's \enquote{waterfall}~\cite{ACL2} and PVS's various proof search strategies~\cite{PVS} have special support for forward reasoning.
Both use techniques similar to Aesop's naive algorithm; in particular, the context of each goal is saturated independently.
ACL2 additionally allows forward rules to be annotated with \emph{trigger terms}, which cause the rules to be run if the trigger terms are matched anywhere in a goal.
Aesop implements a similar mechanism, \emph{rule patterns}, not described here due to space constraints.

>>>>

In many proof assistants, simple \enquote*{white-box} proof automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's~\cite{Isabelle} and Rocq's~\cite{Coq} auto~\cite{IsabelleAuto}, ACL2's \enquote{waterfall}~\cite{ACL2}, various PVS~\cite{PVS-tutorial,PVS} tactics, Lean's~\cite{Lean4} Aesop~\cite{Aesop}, etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.
A typical rule is the $∧$-introduction lemma $A → B → A ∧ B$, which, when applied backwards, reduces a goal $Γ ⊢ A ∧ B$ to subgoals $Γ ⊢ A$ and $Γ ⊢ B$.

@@

\para{White-Box Tactics.}
Among white-box proof automation tactics, to our knowledge only ACL2's \enquote{waterfall}~\cite{ACL2} and PVS's various proof search strategies~\cite{PVS} have special support for forward reasoning.
Both use approaches similar to Aesop's naive algorithm; in particular, the context of each goal is saturated independently.
ACL2 additionally allows forward rules to be annotated with \emph{trigger terms}, which cause the rules to be run if the trigger terms are matched anywhere in a goal.
Aesop implements a similar mechanism, \emph{rule patterns}, not described here due to space constraints.
